The \project{unRAVE} catalog

### 2016-09-02 article/unrave.tex

\subsection{Three-label models ($\teff$, $\logg$, $\feh$)}
\label{sec:a-simple-model}

The model we have introduced is generic in the sense that there is no description for how
complex the vectorizing function $\vecv(\ell)$ should be, or how many $K$ labels are used.
Given previous successes using a second-order polynomial model 
\citep{Ness_2015,Ness_2016,Ho_2016,Casey_2016b}, we will first adopt a three-label ($\teff$, $\logg$, [Fe/H]) model for all (giant and main-sequence) stars. Thus, for our first set of
model experiments,
\begin{eqnarray}\label{eq:vectorizer-three-label}
\vecv(\ell_n) & \rightarrow & \left[1, T_{{\rm eff},n}, \logg_n, [{\rm Fe/H}]_n, T_{{\rm eff},n}^2, \logg_n\cdot{}T_{{\rm eff},n}, \feh_n\cdot{}T_{{\rm eff},n}, \logg_n^2, \feh_n\cdot\logg_n, \feh_n^2\right]
\quad ,
\end{eqnarray}

@@

\subsubsection{A 3-label model for the main-sequence}
\label{sec:the-ms-model}

@@

We made no \emph{a priori} judgements on whether a star was likely a giant
or a dwarf star; we inferred labels for all \rave\ spectra using this 
main-sequence model.  The resulting labels for the survey sample are shown
in Figure \ref{fig:hrd-from-individual-models}.  Significant structure is 
visible.  The increased density of Solar-type stars is consistent with \rave\
observing stars in the local neighbourhood, and the high number of turnoff
and main-sequence stars relative to the subgiant branch is expected from 
the relative lifetimes of these evolutionary phases.  An overdensity
of stars near the giant branch clump is visible, which is likely due to
giant stars existing in the test set, but not in the training set.  In
later sections we will describe how our results from multiple models were
compiled to produce a single catalog.

\subsubsection{A 3-label model for the main-sequence and giant branch}
\label{sec:the-joint-model}

% Chose the regularization hyperparameter.

@@

The giant stars in our training set have stellar parameters and up to 15 
individual abundances from the \apogee\ \aspcap.  Only a subset of these
elements have atomic transitions in the \rave\ wavelength region: 
\ion{O}{1}, \ion{Mg}{1}, \ion{Al}{1}, \ion{Si}{1}, \ion{Ca}{2}, \ion{Ti}{1}, 
\ion{Fe}{1}, and \ion{Ni}{1}.  We excluded [Ti/H] from our abundance list 
because of systematics in the training set \citep{Holtzman_2015,Hawkins_2016}.
Therefore we are left with nine labels in our giant star model: $\teff$, 
$\logg$, and seven individual abundances.

% ARC asks: Should we discuss any tests done with censoring? 
%  			Do we feel like we need to defend our choice to use regularization
%           over censoring? The key result from the censoring tests that I did
%			is that there are a lot of atomic transitions here, but they are
% 			mostly blended. We *need* those blended lines because if you
%			censor all but the strongest line(s) for individual elements then
% 			you are left with a poor predictor of abundance labels.

We used a second-order (i.e., $\teff^2$ is the highest term) polynomial 
vectorizer for the giant star model.  We set the regularization hyperparameter
by performing a grid search of 30 $\Lambda$ points logarithmically spaced from 
$10^{-3}$ to $10^{3}$.  We performed leave-one-out cross-validation on the
training set for each value of $\Lambda$.  Specifically, we removed one star
from the training set, then trained the model, and inferred labels for the 
star that was removed from the training set.  We repeated this procedure for
every star in the training set (for every model with a different $\Lambda$)
and compared the inferred labels against the training set values.  For
comparison purposes, we also performed cross-validation on an unregularized 
($\Lambda = 0$) model using the same training set.  We calculated the bias
and root-mean-square (RMS) deviation between the input labels and those inferred
by cross-validation for every model.  The biases were negligible across all
models.  In Figure \ref{fig:giant-model-hyperparameter-search} we show the
RMS difference in label recovery compared to the unregularized model as a 
percentage improvement.  The enveloped region represents the minimum and 
maximum changes for any label with respect to the unregularized model.
Depending on the regularization strength $\Lambda$, the RMS in a given label
can decrease by $\approx30$\% with respect to the unregularized case.  At the
point of maximum improvement, near $\Lambda = X.XX$, the RMS in all labels has
decreased by at least $X$\%.  As expected the regularization produces a 
sparser matrix of $\vectheta$, with a $\approx20$\% increase in the fraction
of zero-valued entries, particularly for the cross-terms (Figure 
\ref{fig:giant-model-sparsity}, third panel).  Based on the improvement in 
RMS by cross-validation, we adopt $\Lambda = X.XX$ for our giant star model.

Results from cross-validation on our regularized model imply that the precision
in labels is XXX~K in $\teff$, X.XX~dex in $\logg$, and varies between X.XX~dex
and X.XX~dex depending on the elemental abundance.  We made no prior judgement
on whether a star was likely a giant or not; we used this model (and all other
models) to test all XXX,XXX \rave\ spectra.  Figure \ref{fig:test-set-hrd} (left
panel) shows the inferred $\teff$ and $\logg$ values for all \rave\ spectra with
S/N $\geqslant 50$.  This panel undoubtedly includes dwarf stars that cannot be
appropriately classified by our giant model.  In Section \ref{sec:joining-the-models}
we will heuristically combine these results with the main-sequence star model
described in \ref{sec:main-sequence-model}.

\subsection{Joining the models}\label{sec:joining-the-models}

>>>>

\subsection{A 3-label model ($\teff$, $\logg$, $\feh$) for all stars}
\label{sec:a-simple-model}

The model we have introduced is generic in the sense that there is no description for how
complex the vectorizing function $\vecv(\ell)$ should be, or how many $K$ labels are used.
Given previous successes using a second-order (quadratic) polynomial model 
\citep{Ness_2015,Ness_2016,Ho_2016,Casey_2016b}, we will first adopt a three-label ($\teff$, $\logg$, [Fe/H]) model for all (giant and main-sequence) stars. Thus, for our first set of
model experiments,
\begin{eqnarray}\label{eq:vectorizer-three-label}
\vecv(\ell_n) & \rightarrow & \left[1, T_{{\rm eff},n}, \logg_n, [{\rm Fe/H}]_n, T_{{\rm eff},n}^2, \logg_n\cdot{}T_{{\rm eff},n}, \feh_n\cdot{}T_{{\rm eff},n}, \logg_n^2, \feh_n\cdot\logg_n, \feh_n^2\right]
\quad ,
\end{eqnarray}

@@

\subsection{A 3-label model for unevolved stars}
\label{sec:the-ms-model}

@@

We made no \emph{a priori} judgements on whether a star was likely a giant
or a dwarf star; we inferred labels for all \rave\ spectra using this 
main-sequence model.  The resulting labels for the survey sample are shown
in Figure \ref{fig:hrd-from-individual-models}, where significant structure is 
visible.  The increased density of Solar-type stars is consistent with \rave\
observing stars in the local neighbourhood, and the high number of turnoff
and main-sequence stars relative to the subgiant branch is expected from 
the relative lifetimes of these evolutionary phases.  An overdensity
of stars near the giant branch clump is also present.  This artefact is due 
to having giant stars in the test set, but not in the training set, and the
model is (poorly) extrapolating outside the convex hull of the training set.
In Section \ref{sec:joining-the-models} we describe how results from multiple
models are combined to produce a single catalog.

@@

We only have $\teff$, $\logg$, and [Fe/H] for unevolved stars in our training
set from the \epic\ catalog \citep{Huber_2016}.  However, the giant stars 
have stellar parameters ($\teff$, $\logg$) and up to 15 individual abundances
from the \aspcap\ \citep{aspcap_paper}.  A subset of these elements have atomic
transitions in the \rave\ wavelength region: \ion{O}{1}, \ion{Mg}{1}, 
\ion{Al}{1}, \ion{Si}{1}, \ion{Ca}{2}, \ion{Ti}{1}, \ion{Fe}{1}, and \ion{Ni}{1}.  
We excluded [Ti/H] from our abundance list because of systematics in the training
set \citep{Holtzman_2015,Hawkins_2016}.  Therefore we are left with nine labels 
in our giant star model: $\teff$, $\logg$, and seven individual abundances.

We used a quadratic vectorizer for the giant star model, and set the regularization
hyperparameter $\Lambda$ in the same way described in Section \ref{sec:ms-model}.
The results are shown in Figure \ref{fig:set-giant-model-hyperparameters}, where
again the enveloped region represents the minimum and maximum change for any label
with respect to the unregularized case.  At the point of maximum improvement 
near $\Lambda = X.XX$, the RMS in all nine labels has decreased by at least $X$\%, 
and up to $X$\%.  Here the regularization strength also produces a sparser matrix
of $\vectheta$, with a $\approx20$\% increase in the fraction of zero-valued 
entries, particularly for the cross-terms (Figure \ref{fig:giant-model-sparsity},
third panel).  Based on decreasing RMS deviation in the labels, we adopt 
$\Lambda = X.XX$ for our giant star model.

Results from cross-validation on our regularized model imply that the precision
in labels is XXX~K in $\teff$, X.XX~dex in $\logg$, and X.XX~dex (EL) and 
X.XX~dex (EL) depending on the elemental abundance.  We inferred labels for all
XXX,XXX \rave\ spectra using this model, and these results are summarized in
Figure \ref{fig:test-set-hrd} (middle panel).  The red clump is clearly visible
and in the expected location, without requiring any post-analysis calibration.
Artefacts due to dwarf stars in the test set are also present.

\subsection{Joining the models}
\label{sec:joining-the-models}
